---
title: "Cluster Analysis"
author: "Arnab Aich"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
library(readr)
library(mclust)
library(factoextra)
library(ggplot2)
```

**Mclust :** *Model-based clustering based on parameterized finite Gaussian mixture models. Models are estimated by EM algorithm initialized by hierarchical model-based agglomerative clustering. The optimal model is then selected according to BIC (Based on R documentation).*

# 1.

## a)

-   Import Data

```{r}

x1 <- read_csv("D:/OneDrive - Florida State University/MyFSU_OneDrive/FSU Course Work/5635/Datasets/x1.txt", 
    col_names = FALSE,show_col_types = FALSE)
```

-   K means clustering and plot for Initial Clusters

```{r}
K_mean = kmeans(x1,centers = 2)
fviz_cluster(K_mean,data = x1,show.clust.cent = TRUE,stand = TRUE,geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())
```

-   Final clustering using Multivariate Gaussian models

```{r}
cluster = Mclust(x1,G=2,initialization = list(K_mean$centers)) 
fviz_cluster(cluster,data = x1,show.clust.cent = TRUE,stand = TRUE,geom = "point",
             ellipse.type = "convex",ellipse = FALSE, 
             ggtheme = theme_bw())
rm(cluster)
rm(K_mean)
```

## b)

-   Import Data

```{r}

x2 <- read_csv("D:/OneDrive - Florida State University/MyFSU_OneDrive/FSU Course Work/5635/Datasets/x2.txt", 
    col_names = FALSE,show_col_types = FALSE)
```

-   K-means clustering and plot for Initial Clusters

```{r}
K_mean = kmeans(x2,centers = 2)
fviz_cluster(K_mean,data = x1,show.clust.cent = TRUE,stand = TRUE,geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())
```

-   Final clustering using Multivariate Gaussian models

```{r}
cluster = Mclust(x2,G=2,initialization = list(K_mean$centers)) 
fviz_cluster(cluster,data = x2,show.clust.cent = TRUE,stand = TRUE,geom = "point",
             ellipse.type = "convex",ellipse = FALSE, 
             ggtheme = theme_bw())
rm(cluster)
rm(K_mean)
```

## c)

-   Import Data

```{r}
pts <-  read_csv("D:/OneDrive - Florida State University/MyFSU_OneDrive/FSU Course Work/5635/Datasets/pts.txt", 
    col_names = FALSE,show_col_types = FALSE)
```

-   K-means clustering and plot for Initial Clusters

```{r}
K_mean = kmeans(pts,centers = 3)
fviz_cluster(K_mean,data = pts,show.clust.cent = TRUE,stand = TRUE,geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())
```

-   Final clustering using Multivariate Gaussian models

```{r}
cluster = Mclust(pts,G=3,initialization = list(K_mean$centers)) 
fviz_cluster(cluster,data = pts,show.clust.cent = TRUE,stand = TRUE,geom = "point",
             ellipse.type = "convex",ellipse = FALSE, 
             ggtheme = theme_bw())
rm(cluster)
rm(K_mean)
```

## d)

-   K-means clustering and plot for Initial Clusters

```{r}
K_mean = kmeans(pts,centers = 10)
fviz_cluster(K_mean,data = pts,show.clust.cent = TRUE,stand = TRUE,geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())
```

-   Final clustering using Multivariate Gaussian models

```{r}
cluster = Mclust(pts,G=10,initialization = list(K_mean$centers)) 
fviz_cluster(cluster,data = pts,show.clust.cent = TRUE,stand = TRUE,geom = "point",
             ellipse.type = "convex",ellipse = FALSE, 
             ggtheme = theme_bw())
rm(cluster)
rm(K_mean)
```
